# Final Model
Finally! We are ready for our final model! Now, let's use our test and train data. 
```{r}
load('adoption.RData')
set.seed(454)
adoption_split <- initial_split(adoption, prop = .7)
adoption_train <- training(adoption_split)
adoption_test <- testing(adoption_split)
```


## Model Building 
$$
\text{Let }Y_i\text{ be the adoption speed with k=0,1,2.}
\\y_i|(\beta_{0k},\beta_1,...,\beta_15)\sim Bern(\theta_{ki})\text{ where }\theta_{ki}\text{ is the }P(Y_i\leq k)\text{ for the }i^{th}\text{ animal.}
\\log(\frac{\theta_{ki}}{1-\theta_{ki}})=logit(\hat P(Y\leq k))=\beta_{k0}+\beta_1x_1...+\beta_{10}x_{10}+\beta_{11}x_1*x_10+\beta_{12}x_2*x_5+...+\beta_{16}x_2*x_9
\\\text{ where }x_i\text{ for i from 1 to 10 is the indicator for cat, minor/severe injury,}
\\\text{sterilization, age groups 1-5, and Mix breed.}
\\\beta_{12}\text{ is the coefficient for the interaction term of Type*MixedBreed, and }\beta_{13}\text{ to }
\\\beta_{16}\text{ are the coefficients for the interaction terms of AgeGroup*Sterilization.}
\\\beta_{k0}\sim N(m_{k0},s_{k0}^2)
\\\beta_1\sim N(m_1,s_1^2)
\\...$$

```{r cache=TRUE}
final_model <- stan_polr(AdoptionSpeed_Group ~ Type*MixedBreed+Health+Sterilized*AgeGroup, data =adoption_train, prior=R2(0.5,what=NULL),iter=5000, seed = 454)
```


## Diagnostic Plots
Let's look at some diagnostic plots! The trace plots show that all chains overlap without a pattern. The density plot also shows chains overlapping. The pp_check also shows no abnormal patterns. 
```{r cache=TRUE}
mcmc_trace(final_model)
mcmc_dens_overlay(final_model)
pp_check(final_model)
```

## Posterior Inference
```{r cache=TRUE}
final_model_summary<- summary(final_model)
head(as.data.frame(final_model_summary), -2)
```

```{r cache=TRUE}
set.seed(454)
model_data<-adoption_test %>%
  dplyr::select(AdoptionSpeed_Group,Type,Health,AgeGroup,MixedBreed,Sterilized,Vaccinated) %>%
  na.omit()
final_prediction <- posterior_predict(
  final_model,
  newdata = model_data)
ordinal_accuracy2(final_prediction,model_data)
```

## Discussion
What this model tells us. Which factors are important? What can we do to improve our model (cuteness meter)? 


